STAT 450 Project: Real Estate

### 2020-03-05 doc/RealEstate_Peter_Han.Rmd

---
title: 'STAT 450 Project: Real Estate'
author: "Yuting Wen, 44625168"
date: "3/2/2020"
output: pdf_document
---

@@

```{r process, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
library(readxl)

@@

income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
income_budget_compare <- gather(income_budget_compare, TaxType, Amount, Budget:TaxIncome, factor_key=TRUE)
```

Our client is a real estate data analyst in Metro Vancouver who analyzes the market trend to make a market prediction for his company and future client. He is interested in a good prediction model that can accurately predict future mill rates, given the past data like the past mill rates and the past assessment values, in Metro Vancouver. Data cleaning, exploratory data analysis, and linear models with different kinds of regulazations were used to analyze the relationship between mill rate and other factors and predict future mill rates. Data cleaning is performed to aggregate our data into summary statistics and also reduce the number of data entries. Exploratory analysis has shown that there is a strong correlation between mill rate and municipal budget and strong correlation between mill rate and assessment in some municipalities. The results of exploratory analysis have also shown that there is no clear difference among full linear Regression model with all features (Year, TaxClassCode, Municipalities, assessTotal_pct, tax (budget), landTotal, improvementTotal, numProperties) and reduced linear regression models (Ridge, Lasso and Elastic net).

Our goal is to accurately predict the mill rate in 2020 based on past data from 2016 to 2019, for each of the following 21 sub-regions in the Greater Vancouver area:

- Burnaby 
- Coquitlam
- Delta
- Langley City, 
- Langley Township 
- Maple Ridge
- Maple Ridge Rural
- North Vancouver City
- North Vancouver Dist
- Pitt Meadows
- Port Coquitlam
- Port Moody
- Richmond
- Surrey
- Vancouver
- White Rock
- West Vancouver
- Bowen Island
- Anmore
- Belcarra
- Lions Bay

Correlations between mill rates, assessment values, municipal budgets, land assessment values, improvement assessment values, number of properties, tax class codes, municipalities, and year are performed to see the significance of these factors in our predictive model.

A full linear model with all features, a few reduced linear models with regularzations like the LASSO, Ridge Regression are used to train our predictive model. Once we have finished fitting those models, we will use the mean squared training error to evaluate the performance of our model.

To address these problems, we used past assessment data, which is provided by our client. We obtained information about commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019. There are approximately 2.4 million data entries across these 4 years. We selected variables that we believe were relevant to the property tax prediction. These selected variables are:

- Mill Rates (2016-2019)
- Assessment Total (2016-2019)
- Land Assessment Total (2016-2019)
- Improvement Assessment Total (2016-2019)
- Number of Properties (2016-2019)
- Tax Class Code (01, 05, 06)
- Area Code (21 Municipalities)
- Year (2016-2019)

We also collected external data from the Government of British Columbia which could be associated with the tax rate:

The listing data was cleaned to reduce its dimension. We decided to calculate the sum of total assessment value, total land assessment value, total improvement assessment value, budget, and number of properties for each combination of municipality and their corresponding tax class. The mill rate is unique for each combination of municipality and tax class, we have merged this information into our data for each combination of municipality and tax class.

We did some exploratory analysis to visualize the main characteristics of our dataset and found relationships between mill rate and other factors, before we made any prediction on the future mill rate of Metro Vancouver’s real estate market.

We assume that the government aims to match its budget and its income by adjusting the mill rates, so we have plotted a graph of tax income for three different tax classes (01, 05, and 06) vs. budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct. Vancouver, Richmond, and Burnaby were chosen to illustrate such relationship as they are the most representing city in metro vancouver.

We also made scatterplot of mill rate vs. assessment, mil rate vs. municipal budget, mill rate vs. landTotal (total land assessment value), mill rate vs. improvementTotal (total improvement assessment value), and mill rate vs. numOfProperties (number of properties) to see the correlation between each of the two factors. ANOVA analysis was also performed to see the correlations between mill rate vs. tax class, mill rate vs. year and mill rate vs. municipality.

These exploratory analysis and plots helped us to find trends in data within each of its own categories.

In this study, we first built various models then evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values.

Goodness of fit is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data. R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.

Prediction power measures how well models can predict the future values. In this study, we use mean squared prediction error to compare the prediction performance across all the models. To do that, we will divide the data set into training data, used as building models, and test data, used as computing mean squared prediction error.

### Ordinary Linear Model

The explanatory variables we are interested in are Year, TaxClassCode, Municipalities, assessTotal, tax (budget), landTotal, improvementTotal, numOfProperties. We first explored the full linear model using all the available features; then the linear models with regularzations were performed since they potentially have better prediction power.

There are potential outliers in the data set, it might be helpful to refit the model without the outliers to check if the fit of the model is improved. Futher prediction models will be built here.

### Reduced Ordinary Linear Model (Variable Selection)

A list of significant variables are included in the result of the full model. In the next step, we will refit the model with only the significant variables. Goodness of fit will decline with fewer features, however, the performance of prediction power might be improved.

Except for ordinary linear regression, we are also interested in the performance of Ridge, Lasso, and Elastic Net models as they are also in the linear regression family, but more advanced than ordinary linear regression. In these three models, weights are assigned to features. Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively. Elastic net is a combination of Ridge and Lasso.

In the current report, we are mainly focusing on predicting the mill rate using Linear Model. In the future, we will discuss with our supervisor and try to fit a neural network model if possible (R package: neuralnet or keras). Neural Network has the ability to learn and model non-linear and complex relationships. It is flexible and can be used for both classification and prediction. Since we have a large number of data, it makes Neural Network a very good candidate for our prediction model.

Report on the performance of the Neural Network will be included in the next report.

### Results: Exploratory Analysis

We assume that the government aims to match its budget and its income by adjusting the mill rates, so we plotted the predicted tax income vs municipal budget through time into line-chart to visualize their relationship, to make sure our assumption is correct.

For each year from 2016 to 2019, we predicted the tax income of three class codes (01, 05, 06) in each municipality by multiplying access total and mill rate. Vancoouver, Richmond, and Burnaby were chosen because they were the three most representing municipalities in Metro Vancouver, and the plots showed that the predicted tax amount was positively related to municipal budget.

```{r budget, echo=FALSE, warning=FALSE, message=FALSE}
income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
  geom_line() + 
  ggtitle("Figure 1: Budget and Tax Income v.s. Year of Vancouver")
income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
  geom_line() + 
  ggtitle("Figure 2: Budget and Tax Income v.s Year of Richmond")
income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
  geom_line() + 
  ggtitle("Figure 3: Budget and Tax Income v.s Year of Burnaby")
```

### Correlation Analysis

We made 5 scatter plots to visualize the correlation between mill rate and 5 numerical features (assessment total, municipal budget, land assessment total, improvement assessment total, and the number of properties).

@@

# correlation between mill rate and budget for each municipality
rate_tax_correlation <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(rate, tax))
#rate_tax_correlation
# correlation between mill rate and assessment
rate_assess_correlation2 <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(assessTotal, rate))
#rate_assess_correlation2
names(assessment_aggregate)[1] <- paste("Municipality")

## Correlation for continuous variables
assessment_aggregate %>% 
  ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 4: Correlation: Mill Rates and Total Assessment Values") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_aggregate %>% 
  ggplot(aes(x=tax,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + ggtitle("Figure 5: Correlation: Mill rates and Budgets") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_aggregate %>% 
  ggplot(aes(x=landTotal,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 6: Correlation: Mill Rates and Total Land Assessment Values") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_aggregate %>% 
  ggplot(aes(x=improvementTotal,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 7: Correlation: Mill Rates and Total Improvement Assessment Values") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

assessment_aggregate %>% 
  ggplot(aes(x=propertyCount,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 8: Correlation: Mill Rates and Total Number of Properties") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
```

As we can see:

Figure 4, 6, 7, and 8 show that mill rate is negatively correlated with total assessment values, total land assessment values, total improvement assessment values, total number of properties. Budget is not significantly correlated with mill rate as shown in Figure 5.

Outlier detection:

The plots above also show that Vancouver (the dark pink points) are significantly different from other points and should be regarded as an outlier (futher analysis needed here). Some other municipalities (including Richmond, Surrey and Burnaby) that are also distinct in the scatter plots.

### ANOVA Analysis
```{r anova, echo=FALSE, warning=FALSE, message=FALSE}
## Correlation for categorical variables
TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
#summary(TaxClassCode)

@@

                    Table 1: ANOVA analysis of mill rate and three categorical variables

Feature Name  | Pr(>F) | significance levels
------|--------|------
Tax Class Code | <2e-16 | ***(very significant)
Year | 4.2e-06 | **(significant)
Municipality | 0.0768 | .(not significant)

We used three 3 ANOVA tables to analyze the correlation between mill rate and each of the 3 categorical features (TaxClassCode, Year, Municipalities). The table showed that Tax Class Code and Year have a strong correlation with mill rate while the correlation between Municipality and mill rate is not significant at 5% level.

@@

### Variable Selection

                  Table 2: Significance level of all variables of full linear model

After fitting the ordinary linear regression model with all features, the summary statistics in Table 2 showed that municipality, year, tax class code, total assessment value, and total land assess are significant, which indicated that total improvement assessment value, number of properties, and budget(tax) might not have strong impact for the linear model. Therefore we built an ordinary linear regression reduced model with only significant variables.

### Goodness of Fit

                          Table 3: Goodness of fit of all models

Model | Mutiple R-Squared | Adjusted R_Squared | MSE
------|--------|--------|--------|--------
      Ordinary Linear Regression full model | 0.8874| 0.8707| 1.9843
      Ordinary Linear Regression reduced model | 0.8874| 0.8721| 1.9845
      Ridge Regression model |  0.8868| NA| 1.9896
      LASSO Regression model | 0.8873| NA| 1.9855
      Elastic Net Regression model | 0.8625| NA| 2.2366

The table above is the goodness of fit of five models we fitted. All models have relatively high R squared and adjusted R squared values, which indicates that the above models are able to accurately fit the data. Also there is no significant differences among these models in terms of mean squared error, except for Elastic Net model which have higher MSE. We have not used cross validation to reduce model overfitting. Although the results are relatively similar and have high R_squared across all models, we believe using cross validation there will be a clearer difference between the goodness of fit of each model.

### Prediction Power

                        Table 4: Prediction Power of all models
Model | PMSE 
------|--------
      Ordinary Linear Regression full model | 2.5902
      Ordinary Linear Regression reduced model | 2.5237
      Ridge Regression model | 2.5675
      LASSO Regression model | 2.5280
      Elastic Net Regression model | 2.5486

The table above is the prediction power of five models we fitted. All models have relatively low error (PMSE), while there is also no clear difference among them. Futher analysis is needed here.

## Conclusions

From our exploratory data analysis, We have found that when comparing the correlation between mill rate against features in our models, mill rate is highly correlated with total assessment values, total land assessment values, total improvement assessment values, total number of properties, comparing with budget. Also Vancouver is an outlier in our plot. We would like to further investigate Vancouver as a special case.

We can also see that total assessment value, total land assessment value, tax class code, year and municipality are all significant in our linear regression model. We are also interested in the effect of different tax classes in predicting mill rate. Further analysis and prediction will be performed based on our hypothesis.

The linear regression models were able to accurately predict mill rate with relatively low error, but there is still a very high chance that our model is overfitted. For our next report, we are going to use cross-validation to reduce the effect of overfitting.

@@

There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values based on client information; all properties in the same region, classcode, and year should have a unique class rate

- For entries without mill rate, we found the group that they belong to and assign them mill rate in that group.

### Here is some exceptions found:

- In Langley, 2019, Class 06, mill rate is different between assessment type. After talking with the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.

                          Table 4: Groups that miss mill rate

### Model Fitting

#### Ordinary Linear Regression Full Model
```{r full, echo=TRUE, warning=FALSE, message=FALSE}
linear_full<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
#summary(linear_full)

library(broom)
linear_full_fit<-augment(linear_full)
mse_full <- sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
```

#### Ordinary Linear Regression Reduced Model
```{r reduced, echo=TRUE, warning=FALSE, message=FALSE}
reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
#summary(reduced)

@@

#### Ridge
```{r Ridge, echo=TRUE, warning=FALSE, message=FALSE}
library(glmnet)
library(dummies)
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$Municipality)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)

y<-assessment_aggregate$rate
lambdas <- 10^seq(2, -3, by = -.1)

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min

predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)

@@

predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
ridge_mse <- eval_results(y, predictions_train)
```

#### LASSO
```{r LASSO, echo=TRUE, warning=FALSE, message=FALSE}
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

@@

#### Elastic Net
```{r elastic_net, echo=TRUE, warning=FALSE, message=FALSE}
library(caret)
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
```

#### Prediction Power
```{r Prediction power, echo=TRUE, warning=FALSE, message=FALSE}
set.seed(450)
train_ind<-sample(218,218-50)
train<-assessment_aggregate[train_ind,]
test<-assessment_aggregate[-train_ind,]

@@

>>>>

---
title: 'STAT 450 Project: Real Estate'
author: "Peter Han, 14912166"
date: "3/3/2020"
output:
  html_document:
    df_print: paged
---

@@

Our client is a real estate analyst who analyzes the market trend and make market prediction for his company and future client. Data cleaning, exploratory data analysis were used in this project to analyze the relationship between mill rate and other factors. Data cleaning is performed to aggregate our data into summary statistics and also reduce the number of data entries. Exploratory analysis has shown there is a linear trend in mill rates over the years; it has also shown there is a strong correlation between mill rate and municipal budget and strong correlation between mill rate and assessment in some municipalities.

Our client is only interested in accurate predictive models.The goal is to predict the property tax in 2020, for each of the following 21 sub-regions in the Metro Vancouver area:

- Burnaby
- Coquitlam
- Delta
- Langley - City
- Langley - Township
- Maple Ridge
- Maple Ridge Rural 
- North Vancouver - City
- North Vancouver - Dist
- Pitt Meadows
- Port Coquitlam
- Port Moody
- Richmond
- Surrey
- Vancouver
- White Rock
- West Vancouver
- Bowen Island
- Anmore
- Belcarra
- Lions Bay.

Correlations between mil rate, assessment value (Land and Imporvement) and municipal budget were computed to see the significance of these factors in our predictive model. An ANOVA analysis had also been carried out to see the significance of categorical factor such as Year, Municipality and Tax Class Code. A Simple linear model, a linear model with the interaction term, LASSO, Ridge Regression and Elastic Net were used to train our predictive model. Once we have finished the above tasks, we will use the mean square test error to evaluate the performance of our model.

For exploratory data analysis and building predictive model, we used past property assessment data, which is provided by our client. It contains information about some commercial and residential properties in all cities in Metro Vancouver from 2016 to 2019. We selected variables that we believe were relevant to the property tax prediction. These selected variables are:

- Mill Rates (2016-2019)
- Assessment (2016-2019)
- Tax Class Code (01, 05, 06)
- Area Code (Municipality)
- Year
- Land Total (2016 - 2019)
- Improvement Total (2016 - 2019)
- Number of properties

As the client suggested, we have also collected external data from the Government of British Columbia which could be associated with the tax rate:

We decided to calculate the sum of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, and Budget for each combination of municipality and the tax class to reduce the dimension of our data. The mill rate is unique for each combination of municipality and tax class, so we have added this information into our data for each municipality and its tax classes.

Before we made any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to explore and visualize the main characteristics of our dataset and found relationships between Mill Rate vs. Assessment, Mill Rate vs. Municipal Budget, Mill Rate vs. Land Total, Mill Rate vs. Imrpovement Total and Mill Rate vs. Number of Properties.

The municipal government wants their tax income to match their annual spending in order to balance their bedget. Given this piece of information, we want to find and model the relationship between municipal tax income and municipal budget. We plotted a graph of tax income for three different tax classes (01, 05, and 06) vs. budget through time to see the trend of tax income in different municipalities to make sure our assumption about the data is correct. We chose to illustrate such relationship using scatter plot below for Vancouver, Richmond and Burnaby as they are the most representing city in Metro Vancouver.

Mill rate is mainly affected by assessment and municipal budget, so we have plotted scatter plot of mill rate vs. assessment and mil rate vs. municipal budget to see the correlation between each pair of the two factors. ANOVA analysis is also performed to see the correlation between mill rate vs. tax class, mill rate vs. year and mill rate vs. municipality.

In this study, we built a few linear models and evaluated their performances by goodness of fit and prediction power, that is, how well the model explains the data and how well it can predict future values. The definition of goodness of fit and prediction power is given below.

 - **Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data. R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.

 - **Prediction power** measures how well models can predict the future values. We use mean squared prediction power to compare the prediction performance across all of our fitted linear predictive models. To do that, we will divide the data set into training data - used to build our models, and testing data - used to evaluate the prediction power of our models.

### Linear model

For our linear model, Year, TaxClassCode, Municipalities, Assess Total, Land Assessment Total, Improvement Assessment total, Number of properties and tax (budget) are considered. We first explore the full linear model using all the available features. From our exploratory data analysis, we have found Vancouver, Richmond, Surrey and Burnaby are outliers. Further prediction model will be built based on these findings. We are also intesterd in how different tax class can affect our predictive model, so models based on different tax classes will also be considered in the future. 
For our reduced model, we build our linear predictor using only the significant variables from the full model (See table 4). The reduced model's performence in terms of goodness of fit will decrease with fewer features, however, the performance of prediction power might be improved.

Other than simple linear regression, we are also interested in the performance of Ridge Regression, Lasso, and Elastic Net. They are more advanced than simple linear regression as they reduce the variance of the model. In these three models, weights are assigned to features. Ridge regression and Lasso take penalty in sum of absolute values and sum of squared absolute values of weights respectively. Elastic net is a combination of Ridge and Lasso.

In the current report, we are mainly focusing on predicting the mill rate using Linear Model. In the future, we will discuss with our supervisor and try to fit a neural network model if possible (R package: neuralnet or keras). Neural Network has the ability to learn and model non-linear and complex relationships. It is flexible and can be used for both classification and prediction. Since we have a large number of data, it makes Neural Network a very good candidate for our predictive model.

Report on the performance of the Neural Network will be included in the next report.
\pagebreak

###Exploratory Analysis

We assume that the government aims to match its budget and its income by adjusting the mill rates

```{r budget, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
library(readxl)

@@

income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
income_budget_compare <- gather(income_budget_compare, TaxType, Amount, Budget:TaxIncome, factor_key=TRUE)

compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 1: Budget and Tax Income v.s. Year of Vancouver") 
compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 2: Budget and Tax Income v.s. Year of Richmond") 
compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 3: Budget and Tax Income v.s Year of Burnaby")
compare_plot1
compare_plot2
compare_plot3
```

We calculated the total tax income of each municipality by summing the product of assessment total and mill rate of all tax class. Then we plotted the computed tax income vs municipal budget into line-chart to visualize their relationship. We chose Vancoouver, Richmond and Burnaby as representative because they are the three largest municipalities in Metro Vancouver. The plots showed that the computed tax amount was positively and approximatly linearly related to municipal budget through time.

@@

# correlation between mill rate and budget for each municipality
rate_tax_correlation <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(rate, tax))
#rate_tax_correlation
# correlation between mill rate and assessment
rate_assess_correlation2 <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(assessTotal, rate))
#rate_assess_correlation2
names(assessment_aggregate)[1] <- paste("Municipality")
## Correlation for continuous variables
assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 4: Correlations: Mill Rates vs. Total Assessment Values") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 5: Correlations: Mill rates vs. Budgets") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 6: Correlations: Mill Rates vs. Total Assessment Values of Land") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 7: Correlations: Mill Rates vs. Total Assessment Values of Improvement") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 8: Correlations: Mill Rates vs. Total Number of Properties") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))
```

### Correlation Analysis
```{r assess, echo=FALSE}
assess_class
tax_class
land_class
improvementTotal_class
propertyCount_class
```

The above scatter plots showed the correlation between each pair of mill rate vs. assessment total, municipal budget, land assessment total, improvement assessment total, and the number of properties. From all the scatter plots above we have noticed that Vancouver (the dark pink points) is significantly different from other points and should be treated as an outlier/special case. Richmond, Surrey and Burnaby are also distinct in the scatter plots and should be treated as outliers. We believe these municiplaties also have impacts on the correlation between mill rate and its features. Further analysis and prediction model will be used to explore and explain such phenomenon.

### ANOVA Analysis

```{r anova, echo=FALSE, warning=FALSE, message=FALSE}
## Correlation for categorical variables
TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
#summary(TaxClassCode)

@@

**Table 1: Result from ANOVA Analysis**

Feature Name  | Pr(>F) 
------|--------
TaxClassCode | <2e-16 
Year | 4.2e-06 
Municipality | 0.0768

We used three 3 ANOVA analysis to analyze the correlation between mill rate and each of the 3 categorical features - TaxClassCode, Year, Municipalities. TaxClassCode and Year showed a strong correlation with mill rate at 5% significant lete, yet the correlation between Municipality and mill rate is not significant at 5% level.

@@

###Linear Model

**Table 2: Result from Model Used**

Model | Mutiple R-Squared | Adjusted R_Squared | MSE | PMSE 
------|--------|--------|--------|--------
      OLR full | 0.8874| 0.8707| 1.9843| 2.5902
      OLR reduced| 0.8874| 0.8721| 1.9845| 2.5237
      Ridge|  0.8868| NA| 1.9896| 2.5675
      LASSO| 0.8873| NA| 1.9855| 2.5280
      Elastic Net| 0.8625| NA| 2.2366| 2.5486

Above is the goodness of fit of five models. Except for the Elastic Net model which has higher MSE and PMSE, there are no significant differences between these models. We have not used cross-validation to reduce model overfitting.
Although the results are relatively similar and have high R_squared across all models, we believe after using cross-validation there will be a clearer difference between the goodness of fit of each model.

For a dtailed look of how the models were fitted and evaluated, please refer to the appendix.

## Conclusion

From our exploratory data analysis, we can see that assessment total, land total, tax code, year and municipalities are all significant in our model. The client suggested that we can transform numerical factors such as assessment total into percentage change. This data transformation does not perform as well as we expected, which only yields an R^2 of around 0.2 across all of our fitted models. We believe the method that the client has suggested might leave out some important information about the housing market in each municipality. We have also found that when comparing the correlation between mill rate against features in our models, Vancouver is an outlier. We would like to further investigate Vancouver as a special case. We are also interested in the effect of different tax classes in predicting mill rate. Further analysis and prediction will be performed based on our hypothesis.

All fitted linear model were able to make accurate prediction based on means squared error, but there is still a very high chance that our model is overfitted. For our next report, we are going to use cross-validation to reduce the effect of overfitting.

@@

### Code of Linear Regressoin Model
```{r full, warning=FALSE, message=FALSE}
linear_full<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
# summary(linear_full)

library(broom)

linear_full_fit<-augment(linear_full)
mse_full <- sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
```
```{r reduced, echo=FALSE, warning=FALSE, message=FALSE}
reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
#summary(reduced)

@@

```{r Ridges, warning=FALSE, message=FALSE}
library(glmnet)
library(dummies)
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$Municipality)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)

y<-assessment_aggregate$rate
lambdas <- 10^seq(2, -3, by = -.1)
#dim(x)

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
#optimal_lambda
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)

@@

predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
ridge_mse <- eval_results(y, predictions_train)

```

```{r LASSO, warning=FALSE, message=FALSE}
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

@@

```{r elastic_net, echo=FALSE, warning=FALSE, message=FALSE}
library(caret)
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
```

```{r Prediction power, warning=FALSE, message=FALSE}
set.seed(450)
train_ind<-sample(218,218-50)
train<-assessment_aggregate[train_ind,]
test<-assessment_aggregate[-train_ind,]

@@

There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values
Based on client information, all properties in the same region, classcode, and year should have a unique class rate

- For entries without mill rate, we found the group they belong to and assign them mill rate in that group.

###Here is some exceptions found:

- In Langley, 2019, Class 06, mill rate is different between assessment type. After talking to the client, the mill rate for assessment type "land" is regarded as the overall mill rate in that group.

**Table 3: Data with missing mill rate**

###Here is the summary statistics from the full linear regression model

**Table 4: Summary statistics of the full linear regression model**
